Trends of SDR in Thessaloniki

### 2023-06-05 Article.Rmd
Commit 2023-06-02 09:17
Commit 2023-06-05 14:17
Commit 2023-06-05 15:17

The shortwave downward solar irradiance (SDR) at the Earth's surface plays s significant role, on the climate.
Changes of the SDR can be related to major changes on the factors affecting the radiative forcing balance.

In this study we examine the trends of SDR with the ground-based measurements at Thessaloniki for the period 1993 - 2023.

[@Bais2013]

<!-- time span -->
The global radiation was obtained from the pyranometer CM-21, for the period from
`r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
In some algorithms we auxiliary used direct radiation data for comparisons, obtained by a CM-21 pyrheliometer, with data span from 
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.

@@

@@

Due to a significant measurement uncertainty near the horizon, we exclude
all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
Moreover, due to some obstructions around the site (hills and buildings), we exclude data with Azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
At those cases the Sun systematically, is not visible by the instruments.

>>>>

The shortwave downward solar irradiance (SDR) at the Earth's surface plays s significant role, on the climate.
Changes of the SDR can be related to major changes on the factors affecting the radiative forcing balance.
Multiple studies attempt to evaluate the phenomenon and the potential causes [@Ohvril2009; @Schwarz2020; @Wild2009; @Wild2012; @Xia2007; @Zerefos2009].
Although, there are regional variation in the radiation, with both positive and negative trends, most of the studies agree that a major factor is the interaction of aerosols and clouds.

In this study we examine the trends of SDR with ground-based measurements at Thessaloniki for the period 1993 - 2023, as derived from a CM-21 pyranometer.
We reevaluate and extend the dataset used by @Bais2013, applying a different algorithm for the identification of clear-/cloud-sky instances, and we derive the radiation trends for this period under different sky conditions (all-sky, clear-sky and cloud-sky).

....describe CS and cloud definition....

@@

<!-- time span -->
The SDR is equivalent to the whole sky global radiation (GHI), and was obtained with a horizontal leveled, CM-21 pyranometer.
Data spans the period 
`r min(ALL_1_daily_DESEAS$Date)` to `r max(ALL_1_daily_DESEAS$Date)`.
In some algorithms we used as auxiliary data, the direct beam radiation (DNI) for comparisons, obtained by a CHP-1 pyrheliometer, with data availability from 
`r min(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])` to `r max(ALL_1_daily_DESEAS[!is.na(DIR_att), Date])`.

@@

Due to a significant measurement uncertainty near the horizon, we exclude
all measurements with SZA greater than $`r 90-MIN_ELEVA`^\circ$. 
Moreover, due to some obstructions around the site (hills and buildings), we exclude data with Azimuth angle between 
$35^\circ$ and $120^\circ$ and SZA greater than $80^\circ$.
At those instances, Sun is systematically, not visible by the instruments location.
